---
title: "activityMonitoring"
author: "Joabe Leonard"
date: "2 de junho de 2017"
output: html_document
---

1 Code for reading in the dataset and/or processing the data

```{r}
monitoring_data <- read.csv(file='activity.csv',head=TRUE,sep = ",")

summary(monitoring_data)

```

2 Histogram of the total number of steps taken each day

```{r}
steps_by_day <- tapply(monitoring_data$steps, monitoring_data$date, FUN = sum, na.rm = TRUE)

hist(steps_by_day, xlab = "Total number of steps", main="Number of steps by day")
```

3 Mean and median number of steps taken each day


```{r}
mean(steps_by_day, na.rm = TRUE)
median(steps_by_day, na.rm = TRUE)
```

4 Time series plot of the average number of steps taken

```{r}
steps_by_interval <- setNames(aggregate(monitoring_data$steps ~ monitoring_data$interval, monitoring_data$data, mean), c("interval","steps"))

plot(steps_by_interval$interval, steps_by_interval$steps, type='l', 
     main="Average number of steps over all days", xlab="Interval", 
     ylab="Average number of steps")
```

5 The 5-minute interval that, on average, contains the maximum number of steps

```{r}
max(steps_by_interval$steps)
```

6 Code to describe and show a strategy for imputing missing data

```{r}
missing <- is.na(monitoring_data$steps)

table(missing)

fill.value <- function(steps, interval) {
  filled <- NA
  if (!is.na(steps)) 
    filled <- c(steps) else filled <- (steps_by_interval[steps_by_interval$interval == interval, "steps"])
    return(filled)
}
filled.data <- monitoring_data
filled.data$steps <- mapply(fill.value, filled.data$steps, filled.data$interval)
total.steps <- tapply(filled.data$steps, filled.data$date, FUN = sum)


```

7 Histogram of the total number of steps taken each day after missing values are imputed

```{r}
hist(steps_by_day, xlab = "Total number of steps", main="Number of steps by day")

```

8 Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
library(ggplot2)

monitoring_data$date <- as.Date(monitoring_data$date, "%Y-%m-%d")

monitoring_data$day <- weekdays(monitoring_data$date)

monitoring_data$day_type <- c("weekday")

for (i in 1:nrow(monitoring_data)){
  if (monitoring_data$day[i] == "sábado" || monitoring_data$day[i] == "domingo"){
    monitoring_data$day_type[i] <- "weekend"
  }
}

monitoring_data$day_type <- as.factor(monitoring_data$day_type)

table_interval_steps_imputed <- aggregate(steps ~ interval+day_type, monitoring_data, mean)


qplot(interval, steps, data=table_interval_steps_imputed, geom=c("line"), xlab="Interval", 
      ylab="Number of steps", main="") + facet_wrap(~ day_type, ncol=1)

```
